---
title: "Plugging in new calibration algorithms in airGR"
author: "François Bourgin, Guillaume Thirel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Plugging in new calibration algorithms}
  %\VignetteEncoding{UTF-8}
---



```{r setup, warning=FALSE, include=FALSE, fig.keep='none', results='hide'}
library(airGR)
library(DEoptim)
# library(hydroPSO) # Needs R version >= 3.6 or latticeExtra <= 0.6-28 on R 3.5. Archived on 2023-10-16 as requires archived packages 'hydroTSM' and 'hydroGOF'.
library(Rmalschains)
library(caRamel)
library(ggplot2)
library(GGally)
# source("airGR.R")
set.seed(321)
load(system.file("vignettesData/vignetteParamOptim.rda", package = "airGR"))
load(system.file("vignettesData/vignetteParamOptimCaramel.rda", package = "airGR"))
```



# Introduction

## Scope

The Michel's calibration strategy (`Calibration_Michel()` function) is the calibration algorithm proposed in **airGR**. However, other optimization methods can be used in combination with **airGR**.
We show here how to use different R packages to perform parameter estimation.

In this vignette, we use the **GR4J** model to illustrate the different optimization strategies.
In particular, we assume that the R global environment contains input climate data, observed discharge and functions from the [Get Started](V01_get_started.html) vignette, as shown below.
Please note that the calibration period is defined in the `CreateRunOptions()` function .

<!-- ```{r, warning=FALSE, fig.keep='none', results='hide', fig.height=10, fig.width=10, eval=TRUE, echo=FALSE, message=FALSE} -->
<!-- example("Calibration_Michel", echo = FALSE, ask = FALSE) -->
<!-- ``` -->

```{r Calibration_Michel, echo=TRUE, eval=FALSE}
example("Calibration_Michel")
```

In order for the `RunModel_*()` functions to run faster during the parameter estimation process, it is recommended that the outputs contain only the simulated flows (see the `Outputs_Sim` argument in the `CreateRunOptions()` help page).

```{r RunOptions, results='hide', eval=FALSE}
RunOptions <- airGR::CreateRunOptions(FUN_MOD = RunModel_GR4J, InputsModel = InputsModel,
                                      IndPeriod_Run = Ind_Run,
                                      Outputs_Sim = "Qsim")
```

Regarding the different optimization strategies presented here, we refer to each package for in-depth information about the description of the methods used.

Please note that this vignette is only for illustration purposes and does not provide any guidance about which optimization strategies is recommended for the family of the **GR** models.


## Definition of the necessary function

Parameter estimation can be performed by defining a function that takes a parameter set as input and returns the value of the performance criterion.
There are two important steps: the transformation of parameters to real space and the computation of the value of the performance criterion.
Here we choose to minimize the root mean square error.

The change of the repository from the "real" parameter space to a "transformed" space ensures homogeneity of displacement in the different dimensions of the parameter space during the step-by-step procedure of the calibration algorithm of the model.

```{r OptimGR4J, warning=FALSE, results='hide', eval=FALSE}
OptimGR4J <- function(ParamOptim) {
  ## Transformation of the parameter set to real space
  RawParamOptim <- airGR::TransfoParam_GR4J(ParamIn = ParamOptim,
                                            Direction = "TR")
  ## Simulation given a parameter set
  OutputsModel <- airGR::RunModel_GR4J(InputsModel = InputsModel,
                                       RunOptions = RunOptions,
                                       Param = RawParamOptim)
  ## Computation of the value of the performance criteria
  OutputsCrit <- airGR::ErrorCrit_RMSE(InputsCrit = InputsCrit,
                                       OutputsModel = OutputsModel,
                                       verbose = FALSE)
  return(OutputsCrit$CritValue)
}
```


In addition, we need to define the lower and upper bounds of the four **GR4J** parameters in the transformed parameter space:

```{r boundsGR4J, warning=FALSE, results='hide', eval=FALSE}
lowerGR4J <- rep(-9.99, times = 4)
upperGR4J <- rep(+9.99, times = 4)
```



# Local optimization

We start with a local optimization strategy by using the PORT routines (using the `nlminb()` of the `stats` package) and by setting a starting point in the transformed parameter space:

```{r local1, warning=FALSE, results='hide', eval=FALSE}
startGR4J <- c(4.1, 3.9, -0.9, -8.7)
optPORT <- stats::nlminb(start = startGR4J,
                         objective = OptimGR4J,
                         lower = lowerGR4J, upper = upperGR4J,
                         control = list(trace = 1))
```

The RMSE value reaches a local minimum value after 35 iterations.

We can also try a multi-start approach to test the consistency of the local optimization.
Here we use the same grid used for the filtering step of the Michel's calibration strategy (`Calibration_Michel()` function).
For each starting point, a local optimization is performed.

```{r local2, warning=FALSE, results='hide', eval=FALSE}
startGR4JDistrib <- TransfoParam_GR4J(ParamIn = CalibOptions$StartParamDistrib,
                                      Direction = "RT")
startGR4J <- expand.grid(data.frame(startGR4JDistrib))
optPORT_ <- function(x) {
  opt <- stats::nlminb(start = x,
                       objective = OptimGR4J,
                       lower = lowerGR4J, upper = upperGR4J,
                       control = list(trace = 1))
}
listOptPORT <- apply(startGR4J, MARGIN = 1, FUN = optPORT_)
```

We can then extract the best parameter sets and the value of the performance criteria:

```{r local3, warning=FALSE, results='hide', eval=FALSE}
parPORT <- t(sapply(listOptPORT, function(x) x$par))
objPORT <- sapply(listOptPORT, function(x) x$objective)
resPORT <- data.frame(parPORT, RMSE = objPORT)
```

As can be seen below, the optimum performance criterion values (column *objective*) can differ from the global optimum value in many cases, resulting in various parameter sets.

```{r local4, warning=FALSE}
summary(resPORT)
```

The existence of several local minima illustrates the importance of defining an appropriate starting point or of using a multi-start strategy or a global optimization strategy.



# Global optimization

Global optimization is most often used when facing a complex response surface, with multiple local mimina.
Here we use the following R implementation of some popular strategies:

* [DEoptim: differential evolution](https://cran.r-project.org/package=DEoptim)
* [hydroPSO: particle swarm](https://cran.r-project.org/package=hydroPSO)
* [Rmalschains: memetic algorithms](https://cran.r-project.org/package=Rmalschains)


## Differential Evolution

```{r optDE, warning=FALSE, results='hide', eval=FALSE}
optDE <- DEoptim::DEoptim(fn = OptimGR4J,
                          lower = lowerGR4J, upper = upperGR4J,
                          control = DEoptim::DEoptim.control(NP = 40, trace = 10))
```


## Particle Swarm

```{r hydroPSO1, warning=FALSE, results='hide', message=FALSE, eval=FALSE}
# to install the package temporary removed from CRAN
# Rtools needed (windows : https://cran.r-project.org/bin/windows/Rtools/)
# install.packages("https://cran.r-project.org/src/contrib/Archive/hydroPSO/hydroPSO_0.5-1.tar.gz",
#                  repos = NULL, type = "source", dependencies = TRUE)
```


```{r hydroPSO2, warning=FALSE, results='hide', message=FALSE, eval=FALSE}
optPSO <- hydroPSO::hydroPSO(fn = OptimGR4J,
                             lower = lowerGR4J, upper = upperGR4J,
                             control = list(write2disk = FALSE, verbose = FALSE))
```


## MA-LS-Chains

```{r, warning=FALSE, results='hide', eval=FALSE}
optMALS <- Rmalschains::malschains(fn = OptimGR4J,
                                   lower = lowerGR4J, upper = upperGR4J,
                                   maxEvals = 2000)
```



# Results

As it can be seen in the table below, the four additional optimization strategies tested lead to very close optima.

```{r resGLOB, warning=FALSE, echo=FALSE, eval=FALSE}
resGLOB <- data.frame(Algo = c("airGR", "PORT", "DE", "PSO", "MA-LS"),
                      round(rbind(
                        OutputsCalib$ParamFinalR,
                        airGR::TransfoParam_GR4J(ParamIn = optPORT$par                    , Direction = "TR"),
                        airGR::TransfoParam_GR4J(ParamIn = as.numeric(optDE$optim$bestmem), Direction = "TR"),
                        airGR::TransfoParam_GR4J(ParamIn = as.numeric(optPSO$par)         , Direction = "TR"),
                        airGR::TransfoParam_GR4J(ParamIn = optMALS$sol                    , Direction = "TR")),
                        digits = 3))
```
```{r, warning=FALSE, echo=FALSE}
resGLOB
```

<!-- This is an expected result because the response surface for quadratic performance criteria of the **GR4J** model is generally sufficiently well defined in the transformed parameter space to allow using a local optimization strategy instead of a more time consuming global optimization strategy. -->


# Multiobjective optimization

Multiobjective optimization is used to explore possible trade-offs between different performances criteria.
Here we use the following R implementation of an efficient strategy:

* [caRamel: Automatic Calibration by Evolutionary Multi Objective Algorithm](https://cran.r-project.org/package=caRamel)

Motivated by using the rainfall-runoff model for low flow simulation, we explore the trade-offs between the KGE values obtained without any data transformation and with the inverse transformation.

First, the `OptimGR4J()` function previously used is modified to return two values.

```{r, warning=FALSE, results='hide', eval=FALSE}
InputsCrit_inv <- InputsCrit
InputsCrit_inv$transfo <- "inv"

MOptimGR4J <- function(i) {
  if (algo == "caRamel") {
    ParamOptim <- x[i, ]
  }
  ## Transformation of the parameter set to real space
  RawParamOptim <- airGR::TransfoParam_GR4J(ParamIn = ParamOptim,
                                            Direction = "TR")
  ## Simulation given a parameter set
  OutputsModel <- airGR::RunModel_GR4J(InputsModel = InputsModel,
                                       RunOptions = RunOptions,
                                       Param = RawParamOptim)
  ## Computation of the value of the performance criteria
  OutputsCrit1 <- airGR::ErrorCrit_KGE(InputsCrit = InputsCrit,
                                       OutputsModel = OutputsModel,
                                       verbose = FALSE)
  ## Computation of the value of the performance criteria
  OutputsCrit2 <- airGR::ErrorCrit_KGE(InputsCrit = InputsCrit_inv,
                                       OutputsModel = OutputsModel,
                                       verbose = FALSE)
  return(c(OutputsCrit1$CritValue, OutputsCrit2$CritValue))
}
```


## caRamel

caRamel is a multiobjective evolutionary algorithm combining the MEAS algorithm and the NGSA-II algorithm.

```{r, warning=FALSE, results='hide', eval=FALSE}
algo <- "caRamel"
optMO <- caRamel::caRamel(nobj = 2,
                          nvar = 4,
                          minmax = rep(TRUE, 2),
                          bounds = matrix(c(lowerGR4J, upperGR4J), ncol = 2),
                          func = MOptimGR4J,
                          popsize = 100,
                          archsize = 100,
                          maxrun = 15000,
                          prec = rep(1.e-3, 2),
                          carallel = FALSE,
                          graph = FALSE)
```

The algorithm returns parameter sets that describe the pareto front, illustrating the trade-off between overall good performance and good performance for low flow.

```{r, fig.width=6, fig.height=6, warning=FALSE}
ggplot() +
  geom_point(aes(optMO$objectives[, 1], optMO$objectives[, 2])) +
  coord_equal(xlim = c(0.4, 0.9), ylim = c(0.4, 0.9)) +
  xlab("KGE") + ylab("KGE [1/Q]") +
  theme_bw()
```

The parameter sets can be viewed in the parameter space, illustrating different populations.

```{r fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
param_optMO <- apply(optMO$parameters, MARGIN = 1, FUN = function(x) {
  airGR::TransfoParam_GR4J(x, Direction = "TR")
  })
GGally::ggpairs(data.frame(t(param_optMO)), diag = NULL) + theme_bw()
```

```{r fig.height=6, fig.width=12, message=FALSE, warning=FALSE}
RunOptions$Outputs_Sim <- "Qsim"
run_optMO <- apply(optMO$parameters, MARGIN = 1, FUN = function(x) {
  airGR::RunModel_GR4J(InputsModel = InputsModel,
                       RunOptions = RunOptions,
                       Param = x)
  }$Qsim)
run_optMO <- data.frame(run_optMO)

ggplot() +
  geom_line(aes(x = as.POSIXct(InputsModel$DatesR[Ind_Run]),
                y = run_optMO$X1)) +
  geom_line(aes(x = as.POSIXct(InputsModel$DatesR[Ind_Run]),
                y = run_optMO$X54),
            colour = "darkred") +
  scale_x_datetime(limits = c(as.POSIXct("1998-01-01"), NA)) +
  ylab("Discharge [mm/d]") + xlab("Date") +
  scale_y_sqrt() +
  theme_bw()
```
